A distributed key-value store using Raft

### 2014-06-05 proj/paper.tex
paper section 3

However, if a fault occurs during this process which leaves the leader unable to contact a majority of nodes, no response may be generated. To handle this case, the client nodes associate a $600$ms timeout with each request, and will retry the request if this timeout expires. The value of the timeout was chosen to be small while still being long enough to wait out leadership changes. After some number of retries (I picked $5$ to allow for two cycles of timeout-redirects), the client will return an error to the broker. 
\subsection{SET requests}
The beginning and end of a \verb|SET| request look much like those of \verb|GET| requests, wherein the broker submits a request to the client, the client repackages it for the cluster, and the response follows the same chain backwards. The middle part of a \verb|SET| request differs, as setting a value requires a change to be committed into the replicated state. To do this, the leader appends an entry into its log containing the key, the value, the request id, and the client's name. It then runs through the Raft consensus algorithm until the entry is committed. In this implementation, committing a \verb|SET| request entry on any node involves writing it to the replicated state, then messaging the client recorded in the entry to let it know that the request has been completed. This results in all nodes sending the same success message to the client, which improves the odds that the client will see one of them in the event of a network partition. However, it also means that the client has to deal with filtering out duplicates. It does this by using the unique request id, which is included in all messages sent from and to the client.

@@

Later, at line 235, the network partition is resolved, and \verb|A1| is able to contact other nodes again. Immediately, it sends out a heartbeat and is rejected by \verb|A5| because its term is lagging (237). \verb|A1| then receives AppendEntries from the current leader \verb|A3|, rejects them a few times because it is too out-of-date, and finally succeeds in appending 3 entries (241, 249, 253). Meanwhile, \verb|A3| handles the \verb|GET| request from the client in routine fashion, without consulting with the out-of-date \verb|A1|.
\end{document}

>>>>

However, if a fault occurs during this process which leaves the leader unable to contact a majority of nodes, no response may be generated. To handle this case, the client nodes associate a $600$ms timeout with each request, and will retry the request if this timeout expires. If the timeout expires, the client will pick a node at random to contact, because the assumption is that the timeout is due to the leader having gone down in some fashion. The value of the timeout was chosen to be small while still being long enough to wait out leadership changes. After some number of retries (I picked $5$ to allow for two cycles of timeout-redirects), the client will return an error to the broker. 
\subsection{SET requests}
The beginning and end of a \verb|SET| request look much like those of \verb|GET| requests, wherein the broker submits a request to the client, the client repackages it for the cluster, and the response follows the same chain backwards. The middle part of a \verb|SET| request differs, as setting a value requires a change to be committed into the replicated state. To do this, the leader appends an entry into its log containing the key, the value, the request id, and the client's name. It then runs through the Raft consensus algorithm until the entry is committed. In this implementation, committing a \verb|SET| request entry on any node involves writing it to the replicated state, then messaging the client recorded in the entry to let it know that the request has been completed. This results in all nodes sending the same success message to the client, which improves the odds that the client will see one of them in the event of a network partition. However, it also means that the client has to deal with filtering out duplicates. It does this by using the unique request id, which is included in all messages sent from and to the client.

@@

Later, at line 235, the network partition is resolved, and \verb|A1| is able to contact other nodes again. Immediately, it sends out a heartbeat and is rejected by \verb|A5| because its term is lagging (237). \verb|A1| then receives AppendEntries from the current leader \verb|A3|, rejects them a few times because it is too out-of-date, and finally succeeds in appending 3 entries (241, 249, 253). Meanwhile, \verb|A3| handles the \verb|GET| request from the client in routine fashion, without consulting with the out-of-date \verb|A1|.

This script demonstrates the failstop tolerance of the data store, as well as its ability to handle rejoining nodes gracefully. Since a network partition of one node is equivalent to the failstop of that node from the perspective of the rest of the system, it produces the same effect as halting the other node and starting it again.
\subsection{Partitions}
This script is named \verb|leader_partition.chi|, and it runs the scenario where the data store gets partitioned into a group of 3 and a group of 2, where the leader ends up in the partition with two nodes. The desired action by the cluster is for the group of three nodes to recognize that they form a majority, elect a leader among them, and keep going, while the group with two nodes remains ineffective. After the partition is resolved, the two nodes in the smaller group should recognize that they are out-of-date and revert to being followers while updating themselves.

The logging output of a sample run is attached as \verb|partition.out|. The beginning goes much as the previous script, with \verb|A1| being elected leader and processing a \verb|SET| and a \verb|GET| request. At line 105, the network partition is created, and $180$ms later, \verb|A2| times out and begins elections in the group of three. It wins the election at line 125 and becomes the leader of the group of three. Meanwhile, the client has submitted the next \verb|GET| request to \verb|A1| (101), but since it is unable to contact a majority to verify that it is still leader, it is unable to respond. Thus, the client times out, tries to contact \verb|A5| next, and is redirected to \verb|A2| (139, 144). \verb|A2| responds to this immediately (149), and the client proceeds with a \verb|SET| of \verb|k1| to $30$, which runs normally and succeeds (151, 174).

Finally, the partition is resolved at line 175, and the client simultaneously issues a \verb|SET| (177). Next, we see \verb|A1| and \verb|A3| realize that they are behind on terms (181, 185, 189) and revert to being followers of \verb|A2|. At the same time, \verb|A2| handles the \verb|SET| request from the client and distributes the new value of \verb|k2| to all followers, including \verb|A1| and \verb|A3|, and they both catch up at log index $4$. The \verb|SET| proceeds, and after \verb|A1| and \verb|A3| are caught up, the cluster reverts to normal operation.

This script demonstrates the partition tolerance of the data store. More specifically, it retains consistency while experiencing a network partition, while suffering brief loss of availability during the leader changeover (with potentially unbounded duration of the changeover process).
\section{Issues and lessons learned}
\end{document}
